---
title: "Parent-wise cross-validation to check the accuracy of predicting cross (co)-variances"
author: "Marnin Wolfe"
date: "2021-May-14"
output: 
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, 
                      eval = FALSE, # <- NOTE THAT EVAL SET TO FALSE!
                      tidy='styler', tidy.opts=list(strict=FALSE,width.cutoff=100), highlight=TRUE)
```

# Previous step

4.  [Preprocess data files](04-PreprocessDataFiles.html): Prepare haplotype and dosage matrices, pedigree and BLUPs, genetic map *and* recombination frequency matrix, for use in predictions.

# Automating cross-validation

In the manuscript, the cross-validation is documented many pages and scripts, [documented here](https://wolfemd.github.io/PredictOutbredCrossVar/).

For ongoing GS, I have a function `runCrossVal()` that manages all inputs and outputs easy to work with pre-computed accuracies.

Goal here is to make a function: `runParentWiseCrossVal()`, or at least make progress towards developing one.

*However*, for computational reasons, I imagine it might still be best to separate the task into a few functions.

My goal is to simplify and integrate into the pipeline used for NextGen Cassava. In the paper, used multi-trait Bayesian ridge-regression (MtBRR) to obtain marker effects, and also stored posterior matrices on disk to later compute posterior mean variances. This was computationally expensive and different from my standard univariate REML approach. I think MtBRR and PMV are probably the least biased way to go... but...

For the sake of testing a simple integration into the in-use pipeline, I want to try univariate REML to get the marker effects, which I'll subsequently use for the cross-validation.

Revised the functions in **`package:predCrossVar`** to increase the computational efficiency. Not yet included into the actual R package but instead sourced from `code/predCrossVar.R`. Additional speed increases were achieved by extra testing to optimize balance of `OMP_NUM_THREADS` setting (multi-core BLAS) and parallel processing of the  crosses-being-predicted. Improvements will benefit users predicting with REML / Bayesian-VPM, but probably worse for Bayesian-PMV.

# Set-up server computing env.

Use a a singularity image from the rocker project, as recommended by Qi Sun to get an OpenBLAS-linked R environment that packages can easily be installed on.

**This first chunk is one-time only and doesn't take long. Saves a 650Mb `*.sif` file to server's /workdir/**
```{bash, eval=F}
# copy the project data
cd /home/jj332_cas/marnin/;
cp -R implementGMSinCassava /home/$USER;
# the project directory can be in my networked folder for 2 reasons:
# 1) singularity will automatically recognize and be able to access it
# 2) My analyses not read/write intensive; don't break server rules/etiquette 
# set up a working directory on the remote machine
mkdir /workdir/$USER
cd /workdir/$USER/; 

# pull a singularity image and save in the file rocker.sif
# next time you use the rocker.sif file to start the container
singularity pull rocker.sif docker://rocker/tidyverse:latest;
```

For analysis, operate each R session within a singularity Linux shell within a screen shell.
```{bash, eval=F}
# 1) start a screen shell
screen; 
# 2) start the singularity Linux shell inside that
singularity shell /workdir/$USER/rocker.sif; 
# Project directory, so R will use as working dir.
cd /home/mw489/implementGMSinCassava/;
# 3) Start R
R
```

# Run parent-wise cross-validation

Fully-tested `runParentWiseCrossVal()` and component functions are in the `code/parentWiseCrossVal.R` script. 

Below, source it and use it for a full cross-validation run. 

```{r install packages if needed}
# install.packages(c("RhpcBLASctl","here","rsample","sommer","psych"))
# install.packages(c("psych"))
```

```{r primary inputs}
require(tidyverse); require(magrittr); 
# 5 threads per Rsession for matrix math (openblas)
RhpcBLASctl::blas_set_num_threads(5)

# SOURCE CORE FUNCTIONS
source(here::here("code","parentWiseCrossVal.R"))
source(here::here("code","predCrossVar.R"))

# PEDIGREE
ped<-read.table(here::here("output","verified_ped.txt"),
                header = T, stringsAsFactors = F) %>% 
  rename(GID=FullSampleName,
         damID=DamID,
         sireID=SireID) %>% 
  dplyr::select(GID,sireID,damID)
# Keep only families with _at least_ 2 offspring
ped %<>%
  semi_join(ped %>% count(sireID,damID) %>% filter(n>1) %>% ungroup())

# BLUPs
blups<-readRDS(file=here::here("data","blups_forCrossVal.rds")) %>% 
  dplyr::select(-varcomp)

# GENOMIC RELATIONSHIP MATRICES (GRMS)
grms<-list(A=readRDS(file=here::here("output","kinship_A_IITA_2021May13.rds")),
           D=readRDS(file=here::here("output",
                                     "kinship_domGenotypic_IITA_2021July5.rds")))
## using A+domGenotypic (instead of domClassic used previously)
## will achieve appropriate dom effects for predicting family mean TGV
## but resulting add effects WILL NOT represent allele sub. effects and thus
## predictions won't equal GEBV, allele sub. effects will be post-computed
## as alpha = a + d(q-p)

# DOSAGE MATRIX
dosages<-readRDS(file=here::here("data",
                                 "dosages_IITA_filtered_2021May13.rds"))

# RECOMBINATION FREQUENCY MATRIX
recombFreqMat<-readRDS(file=here::here("data",
                                       "recombFreqMat_1minus2c_2021May13.rds"))
# HAPLOTYPE MATRIX
## keep only haplos for parents-in-the-pedigree
## those which will be used in prediction, saves memory
haploMat<-readRDS(file=here::here("data","haps_IITA_filtered_2021May13.rds"))
parents<-union(ped$sireID,ped$damID) 
parenthaps<-sort(c(paste0(parents,"_HapA"),
                   paste0(parents,"_HapB")))
haploMat<-haploMat[parenthaps,colnames(recombFreqMat)]

# SELECTION INDEX WEIGHTS
## from IYR+IK
## note that not ALL predicted traits are on index
SIwts<-c(logFYLD=20,
         HI=10,
         DM=15,
         MCMDS=-10,
         logRTNO=12,
         logDYLD=20,
         logTOPYLD=15,
         PLTHT=10) 
```

Server 1: modelType="DirDom"

cbsulm17 - 112 cores, 512 GB RAM
```{r model DirDom - 5 full reps - parent-wise CV}
cvDirDom_5rep5fold<-runParentWiseCrossVal(nrepeats=5,nfolds=5,seed=84,
                                          modelType="DirDom",
                                          ncores=20,
                                          outName="output/cvDirDom_5rep5fold",
                                          ped=ped,
                                          blups=blups,
                                          dosages=dosages,
                                          haploMat=haploMat,
                                          grms=grms,
                                          recombFreqMat = recombFreqMat,
                                          selInd = TRUE, SIwts = SIwts)
saveRDS(cvDirDom_5rep5fold,here::here("output","cvDirDom_5rep5fold_predAccuracy.rds"))
```

Server 2: modelType="AD"

cbsulm29 - 104 cores, 512 GB RAM

```{r model AD - 5 full reps - parent-wise CV}
grmsAD<-list(A=readRDS(file=here::here("output","kinship_A_IITA_2021May13.rds")),
             D=readRDS(file=here::here("output",
                                       "kinship_D_IITA_2021May13.rds")))
rm(grms)
cvAD_5rep5fold<-runParentWiseCrossVal(nrepeats=5,nfolds=5,seed=84,
                                      modelType="AD",
                                      ncores=20,
                                      outName="output/cvAD_5rep5fold",
                                      ped=ped,
                                      blups=blups,
                                      dosages=dosages,
                                      haploMat=haploMat,
                                      grms=grmsAD,
                                      recombFreqMat = recombFreqMat,
                                      selInd = TRUE, SIwts = SIwts)
saveRDS(cvAD_5rep5fold,here::here("output","cvAD_5rep5fold_predAccuracy.rds"))
```

# Next step / Results

See [Results](06-Results.html): Home for plots and summary tables.



