---
title: "Empirical inputs for simulations - IITA"
site: workflowr::wflow_site
date: "2021-Aug-13"
output: 
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, 
                      tidy='styler', tidy.opts=list(strict=FALSE,width.cutoff=100), highlight=TRUE)
```

```{bash set-up remote R environment, eval=F}
# 1) start a screen shell 
screen; # or screen -r if re-attaching...

# cbsurobbins SLURM 
salloc -n 20 --mem=60G --time=06:00:00;

# 2) start the singularity Linux shell inside that
singularity shell ~/rocker2.sif; 
# Project directory, so R will use as working dir.
cd /home/mw489/IITA_2021GS/;
# 3) Start R
R
```

# Estimate selection error

## Concept

Determine error-variance vs. plot-size/rep-number scaling for [input to simulations](https://wolfemd.github.io/IITA_2021GS/BreedingSchemeOpt/).

Our downstream objective is to simulate a baseline and alternative breeding pipelines (variety development pipelines; VDPs) based on proposed differences in plot-size (number of plants), number of reps, locations and overall trial-stage size.

We can't compute the same selection index for each trial because of variation in the traits scored.

Below, I outline an approach that integrates into the existing GS pipeline and deals with heterogeneity in the traits observed from trial-to-trial.

Approach:

-   Use the *SELECTION INDEX* GETGV from genomic prediction using all entire available training population and all of the latest available data as a best estimate of "true" net merit

    -   The predictions documented [here](https://wolfemd.github.io/IITA_2021GS/06-GenomicPredictions.html) is the most up-to-date. See the summary of genomic prediction results [here](https://wolfemd.github.io/IITA_2021GS/07-Results.html#Genomic_Predictions).

    -   Alternative best estimate of "true" net merit: actual national performance trial data (e.g. from NCRPs in Nigeria?)

-   For each trial, analyze the cleaned plot-basis data:

    1.  Fit a univariate mixed-model to each trait scored

    2.  Extract trial-specific BLUPs for whatever clones were present

    3.  Compute the SELIND for the current trial using BLUPs for whatever component traits were scored ($SI_{TrialBLUP}$).

    4.  Regress $SI_{GETGV}$ on the $SI_{TrialBLUP}$

    5.  Extract the $\hat{\sigma}^2_e$ of the regression as the trial-specific estimate of the selection error

-   Using the results from all trials: Regress $\hat{\sigma}^2_e$ on plot-size. Weight by number of clones available to measure $\hat{\sigma}^2_e$. Extract coefficients to generate relative scaling profiles of error variances to input for simulating VDPs.

## Analysis

```{r load getgv and support data}
library(genomicMateSelectR); 
library(tidyverse)

# SELIND GETGVS (for input to estimateSelectionError func below)
gpreds<-readRDS(file = here::here("output","genomicPredictions_full_set_2021Aug09.rds"))
getgvs<-gpreds$gblups[[1]] %>% 
  filter(predOf=="GETGV") %>% 
  select(GID,SELIND)

# CLEANED PLOT-LEVEL TRIAL DATA
dbdata<-readRDS(here::here("output","IITA_ExptDesignsDetected_2021Aug08.rds"))
### Restrict consideration to >2012 
### to measure the selection error during the current "era" at IITA.
trials2keep<-dbdata %>% 
  filter(studyYear>=2013) %>% 
  distinct(studyYear,locationName,studyName,TrialType,CompleteBlocks,IncompleteBlocks,MaxNOHAV) %>% 
  filter(!is.na(MaxNOHAV)) %$% 
  unique(studyName)
length(trials2keep) %>% paste0(.," trials")
# [1] 641 trials
dbdata %>% filter(studyName %in% trials2keep) %>% nrow %>% paste0(.," plots")
# [1] 182195 plots
trialdata<-dbdata %>% 
  filter(studyYear>=2013,
         studyName %in% trials2keep) %>% 
  nest(TrialData=-c(studyYear,locationName,studyName,TrialType,CompleteBlocks,IncompleteBlocks,MaxNOHAV))
trialdata %<>% 
  mutate(propGenotyped=map_dbl(TrialData,
                              ~length(which(!is.na(unique(.$FullSampleName))))/length(unique(.$GID))))

# SELECTION INDEX WEIGHTS
## from IYR+IK
## note that not ALL predicted traits are on index
SIwts<-c(logFYLD=20,
         HI=10,
         DM=15,
         MCMDS=-10,
         logRTNO=12,
         logDYLD=20,
         logTOPYLD=15,
         PLTHT=10) 

# SOURCE FUNCTION estimateSelectionError() 
source(here::here("code","estimateSelectionError.R"))
```

```{r summarize trial data}
trialdata %>% unnest(TrialData) %>% 
  summarise(Nplots=nrow(.),
            across(c(locationName,studyYear,studyName,TrialType,GID), ~length(unique(.)),.names = "N_{.col}")) %>% 
  rmarkdown::paged_table()
```

**Summary of available plot-level data:** See [here](01-cleanTPdata.html) for details about the cleaned trial data most recently downloaded from IITA/Cassavabase and used below. Number of unique plots, locations, years, etc. in the cleaned plot-basis data.

```{r, cols.print=15}
trialdata %>% unnest(TrialData) %>% 
  count(TrialType,CompleteBlocks,IncompleteBlocks) %>% 
  spread(TrialType,n) %>% 
  rmarkdown::paged_table()
```

**Count the trial designs by TrialType:**

```{r plot MaxNOHAV vs. TrialType, cols.print=16}
trialdata %>% unnest(TrialData) %>% 
  distinct(studyYear,locationName,studyName,TrialType,CompleteBlocks,IncompleteBlocks,MaxNOHAV) %>% 
  filter(!is.na(MaxNOHAV)) %>% 
  mutate(TrialType=factor(TrialType,levels=c("CrossingBlock","GeneticGain","CET","ExpCET","PYT","AYT","UYT","NCRP"))) %>% 
  ggplot(.,aes(x=TrialType,y=MaxNOHAV,fill=TrialType)) + 
  geom_boxplot(notch = T) + 
  theme_bw() + theme(axis.text.x = element_text(angle=45,vjust=.5)) +
  labs(title = "Max number harvested as a proxy for planned plot size",
       subtitle="MaxNOHAV = The maximum number stands harvested per trial. studyYear>=2013")
```

```{r unit test inputs for estimateSelectionError, eval=F}
###### unit test inputs for estimateSelectionError
# TrialData<-trialdata$TrialData[[1]]
# CompleteBlocks<-trialdata$CompleteBlocks[[1]]
# IncompleteBlocks<-trialdata$IncompleteBlocks[[1]]
# TrialData<-trialdata %>% filter(propGenotyped>0.75) %>% slice(4) %$% TrialData[[1]]
# CompleteBlocks<-trialdata %>% filter(propGenotyped>0.75) %>% slice(4) %$% CompleteBlocks[[1]]
# IncompleteBlocks<-trialdata %>% filter(propGenotyped>0.75) %>% slice(4) %$% IncompleteBlocks[[1]]
# ncores=4
# rm(TrialData,CompleteBlocks,IncompleteBlocks)
```

Run function `estimateSelectionError()` across trials to estimation selection errors.

```{r, eval=F}
require(furrr); plan(multisession, workers = 20)
options(future.globals.maxSize=+Inf); options(future.rng.onMisuse="ignore")
trialdata %<>% 
  mutate(SelectionError=future_pmap(.,estimateSelectionError,
                                    SIwts=SIwts,getgvs=getgvs))
plan(sequential)
saveRDS(trialdata,here::here("output","estimateSelectionError.rds"))

```

## Result

Out of 633 trials, 516 produced successful model fits and subsequent estimates of TrialMSE (selection index error).

```{r}
library(genomicMateSelectR)
estSelError<-readRDS(here::here("output","estimateSelectionError.rds"))
estSelError %<>% 
  select(-TrialData) %>% 
  unnest(SelectionError) %>% 
  select(-SI_BLUPs,-BLUPs,-SelectionError) %>% 
  filter(!is.na(TrialMSE))
```

Here's the `str()` of the estimates I made.

```{r}
estSelError %>% str
```

-   **`cor2si`** = correlation between SI computed from each trial's BLUPs and the SI computed from GETGV (all training data and traits used)
-   **`r2_si`** = r-squared, regression of SI_GETGV on SI_TrialBLUP
-   **`TrialMSE`** = mean squared error from that regression
-   **`NcloneForReg`** = the number of clones with estimates of both SI_TrialBLUP and SI_GETGV for a given trial. Avoid considering trials with too few data points.

I made two plots below, one of the `cor2si` and one of the `TrialMSE` vs. MaxNOHAV (the proxy for plot size). Trial-results are scaled by NcloneForReg and a smooth spline is fit.
```{r cor2si vs MaxNOHAV}
estSelError %>% 
  ggplot(.,aes(x=MaxNOHAV,y=cor2si,color=NcloneForReg,size=NcloneForReg)) + 
  geom_point() + geom_smooth() + theme_bw() + theme(panel.grid = element_blank()) +
  labs(title = "The cor(SI_GETGV,SI_TrialBLUPs) vs. MaxNOHAV as a proxy for plot size")
```

```{r TrialMSE vs MaxNOHAV}
estSelError %>% 
  ggplot(.,aes(x=MaxNOHAV,y=TrialMSE,color=NcloneForReg,size=NcloneForReg)) + 
  geom_point() + geom_smooth() + theme_bw() + theme(panel.grid = element_blank()) +
  labs(title = "Trial-specific selection error ests. vs. MaxNOHAV")
```

In my opinion, there is no clear pattern. I am biased to seeing a hint of a trend.

Look at a correlation matrix and do some regression analysis below.

Try to measure an effect size of increasing the number of stands per plot.

```{r}
estSelError<-readRDS(here::here("output","estimateSelectionError.rds"))
# extract some covariates which might be predictive of cor2si, 
# which I think is on a uniform scale

# Nclone in trial: want est. of plot-size effect NOT dependent on Nclones
# estSelError$TrialData[[1]] %$% length(unique(GID)) 
# MeanPropNOHAV: measure trial quality by how "full" the plots were at harvest
# estSelError$TrialData[[1]] %$% mean(PropNOHAV) 
estSelError %<>%
  mutate(Nclones=map_dbl(TrialData,~length(unique(.$GID))),
         MeanPropNOHAV=map_dbl(TrialData,~mean(.$PropNOHAV,na.rm=T)),
         Nobs=map_dbl(TrialData,~nrow(.)),
         TrialDesign=paste0("Complete",CompleteBlocks,"Incomplete",IncompleteBlocks)) %>% 
  select(-TrialData) %>% 
  unnest(SelectionError) %>% 
  select(-SI_BLUPs,-BLUPs,-SelectionError) %>% 
  filter(!is.na(TrialMSE)) %>% 
  mutate(propAvailForReg=NcloneForReg/Nclones)
estSelError %>% str
```

```{r}
estSelError %>% 
  select(cor2si,TrialMSE,MaxNOHAV,NcloneForReg,Nclones,MeanPropNOHAV,Nobs,propAvailForReg) %>% 
  cor(., use='pairwise.complete.obs') %>% 
  round(.,2) %>% 
  corrplot::corrplot(type="lower")
```

- `MeanPropNOHAV`: Avg. proportion of stands harvested per plot. Measure of trial quality / success.
- `Nclones` and `Nobs`: Measures of trial size
- `propAvailForReg = NcloneForReg/Nclones`: a weight for the regressions below. The maximum weight is to be on trials with as many clones as intended scored on the selection index. 

Linear model below: include all feasible co-variables _and_ an interaction b/t TrialDesign and MaxNOHAV (plot size)
```{r}
lm(cor2si~MaxNOHAV+TrialDesign+TrialDesign*MaxNOHAV+studyYear+Nclones+MeanPropNOHAV,
   data = estSelError,
   weights = propAvailForReg) %>% summary
```
```{r}
lm(TrialMSE~MaxNOHAV+TrialDesign+TrialDesign*MaxNOHAV+studyYear+Nclones+MeanPropNOHAV,
   data = estSelError,
   weights = propAvailForReg) %>% summary
```
<6% variance explained (`cor2si`) and 10% for `TrialMSE`. 

Next simplify a bit, remove the interactions and TrialDesign effect, but keep the covariates.

Simplify it a bit, e.g.: `cor2si~MaxNOHAV+studyYear+Nclones+MeanPropNOHAV`
```{r}
lm_cor2si<-lm(cor2si~MaxNOHAV+studyYear+Nclones+MeanPropNOHAV,
   data = estSelError,
   weights = propAvailForReg)
lm_cor2si %>% summary
```

```{r}
lm(TrialMSE~MaxNOHAV+studyYear+Nclones+MeanPropNOHAV,
   data = estSelError,
   weights = propAvailForReg) %>% summary
```

```{r}
# VarT=100
# h2=0.2
# VarE=(1-h2)*VarT
# tibble(plotSize=1:50) %>% 
#   mutate(expCorToSI=lm_cor2si$coefficients[["MaxNOHAV"]]*plotSize,
#          sqCorToSI=round(expCorToSI^2,2),
#          VarE=VarE*(1-sqCorToSI))
```

# Conclusions

The analysis seems far from conclusive. Do we even expect the relation b/t plot-size and error variance to be linearly decreasing? 

Certainly this will vary by breeding, location, and much more.

There are also possibly many alternative methods to analyze this problem, and alternative datasets to choose.


**NEW APPROACH:** While inconclusive, this exercise emphasizes a key concern for conducting simulations that alter the VDP: that the cost-benefit balance could depend on the relative information value/selection accuracy/error variance of different plot sizes and trial configurations. As a result, I propose to simulate a range of error-vs-plot size as part of the [baseline simulation](baselineSim.html) of each breeding program. ***If*** we observe a shift-point in the cost-benefit analysis we can then work with breeding programs to determine where their data indicate they lie and what changes are subsequently recommended. 
