---
title: "Empirical inputs for simulations - IITA"
site: workflowr::wflow_site
date: "2021-Aug-13"
output: 
  workflowr::wflow_html:
    code_folding: hide
    toc: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = F, 
                      tidy='styler', tidy.opts=list(strict=FALSE,width.cutoff=100), highlight=TRUE)
```

# Estimate selection error

## Approach 1: direct estimation with multivar mixed model

Try a multivariate model with the proper error and other heterogeneous variances.

This will be for DIRECT calculation, if I can get the models to converge, of the variance components of interest.

Goal is to get estimate of errorCov matrix *per* stage.

Then compute SI errorVar *per* stage.

Things not modeled, which could be input to sims:

-   GxL, GxYr, GxYrxL variance
-   genCor among locs

```{bash, eval=F}
screen;
cd ~/IITA_2021GS/;
salloc -n 8 --mem=60G --time=06:00:00;
export PATH=/programs/R-4.0.5clean-p/bin:$PATH
export OMP_NUM_THREADS=8
R;
```

```{r setup inputs}
library(genomicMateSelectR); 
library(tidyverse)

# CLEANED PLOT-LEVEL TRIAL DATA
dbdata<-readRDS(here::here("output","IITA_ExptDesignsDetected_2021Aug08.rds"))
# SELECTION INDEX WEIGHTS
SIwts<-c(logFYLD=20,
         HI=10,
         DM=15,
         MCMDS=-10,
         logRTNO=12,
         logDYLD=20,
         logTOPYLD=15,
         PLTHT=10) 
# FILTER TRIALS TO-BE-CONSIDERED
### Restrict consideration to >2012 
### to measure the selection error during the current "era" at IITA.
### Only trials with >=50% genotyped and key TrialTypes
### Keep only trials with full plotWidth and plotLength meta-data
### Calc plotArea=plotWidth*plotLength (in meters-squared)
trialdata<-dbdata %>% 
  filter(studyYear>=2013,
         !is.na(MaxNOHAV),
         !is.na(plotWidth),
         !is.na(plotLength),
         !is.na(PropNOHAV)) %>% 
  nest(TrialData=-c(studyName,TrialType,plotWidth,plotLength,CompleteBlocks,IncompleteBlocks,MaxNOHAV)) %>% 
  mutate(propGenotyped=map_dbl(TrialData,
                              ~length(which(!is.na(unique(.$FullSampleName))))/length(unique(.$GID))),
         plotArea=plotWidth*plotLength,
         IncompleteBlocks=ifelse(IncompleteBlocks==TRUE,"Yes","No"),
         CompleteBlocks=ifelse(CompleteBlocks==TRUE,"Yes","No")) %>% 
  filter(propGenotyped>=0.5,
         TrialType %in% c("GeneticGain","CET","ExpCET","PYT","AYT","UYT")) %>%
  unnest(TrialData) %>% 
  select(yearInLoc,studyName,studyYear,locationName,TrialType,
         plotArea,plotWidth,plotLength,
         CompleteBlocks,IncompleteBlocks,observationUnitDbId,
         GID,trialInLocYr,repInTrial,blockInRep,PropNOHAV,MaxNOHAV,
         all_of(names(SIwts)))
trialdata %>% nrow(.) %>% paste0(.," plots"); 
# [1] "108254 plots"
trialdata %>% distinct(studyName) %>% nrow(.) %>% paste0(.," trials");
# [1] "382 trials"
```

Filters applied to the data-to-be-considered:

-   `studyYear>=2013`
-   `propGenotyped>=0.5`
-   `TrialType %in% c("GeneticGain","CET","ExpCET","PYT","AYT","UYT")`
-   Must have plotLength, plotWidth *and* MaxNOHAV meta-data.

It turns out, most IITA trials at least already have this.

```{r plotArea vs. TrialType}
trialdata %>% 
  distinct(studyYear,locationName,studyName,TrialType,CompleteBlocks,IncompleteBlocks,plotArea,MaxNOHAV) %>% 
  mutate(TrialType=factor(TrialType,levels=c("CrossingBlock","GeneticGain","CET","ExpCET","PYT","AYT","UYT","NCRP"))) %>% 
  ggplot(.,aes(x=TrialType,y=plotArea,fill=TrialType)) + 
  geom_boxplot(notch = T) + 
  theme_bw() + theme(axis.text.x = element_text(angle=45,vjust=.5)) +
  labs(title = "Plot Area (m-squared) by TrialType",
       subtitle="plotArea = plotWidth*plotLength. studyYear>=2013")
```

```{r MaxNOHAV vs. TrialType}
trialdata %>% 
  distinct(studyYear,locationName,studyName,TrialType,CompleteBlocks,IncompleteBlocks,plotArea,MaxNOHAV) %>% 
  mutate(TrialType=factor(TrialType,levels=c("CrossingBlock","GeneticGain","CET","ExpCET","PYT","AYT","UYT","NCRP"))) %>% 
  ggplot(.,aes(x=TrialType,y=MaxNOHAV,fill=TrialType)) + 
  geom_boxplot(notch = T) + 
  theme_bw() + theme(axis.text.x = element_text(angle=45,vjust=.5)) +
  labs(title = "Max number harvested as a proxy for planned plot size",
       subtitle="MaxNOHAV = The maximum number stands harvested per trial. studyYear>=2013")
```

Much less clear difference between trials using MaxNOHAV.

I decided to take a narrow view of plot configurations and analyze only trials confi=orming to the common (median) plotArea for each TrialType. Does not exclude too many.

Below, I work carefully up to a multivariate model with 7 of the 8 SELIND traits. Starting with homogenous variances and one trait, then 2 traits, 2 traits + heterogenous error by TrailType, 4 + heterog. error, and finally 7 traits skipping PLTHT b/c of \>50% missingness.

```{r keep trials with 'standard' plotArea}
trialdata %<>% 
  semi_join(trialdata %>% 
              distinct(TrialType,studyName,plotArea,plotWidth,plotLength) %>% 
              group_by(TrialType) %>% 
              summarize(plotArea=median(plotArea)) %>% ungroup())

trialdata %>% nrow(.) %>% paste0(.," plots"); 
# [1] "70301 plots"
trialdata %>% distinct(studyName) %>% nrow(.) %>% paste0(.," trials");
# [1] "240 trials"
```

```{r}
trialdata %>% 
  distinct(TrialType,plotArea) %>% 
  arrange(plotArea)
```

### Work up to the full analysis

```{r one trait, eval=F}
MultiTrialTraitData<-trialdata %>%
  filter(!is.na(DM)) %>% 
  mutate(across(c(GID,yearInLoc,
                  CompleteBlocks,
                  IncompleteBlocks,
                  trialInLocYr,
                  repInTrial,
                  blockInRep),as.factor)) %>% 
  droplevels

fixedFormula="DM ~ yearInLoc"
randFormula=paste0("~idv(GID) + idv(trialInLocYr) + at(CompleteBlocks,'Yes'):repInTrial ",
                   "+ at(IncompleteBlocks,'Yes'):blockInRep")

require(asreml); 
fixedFormula<-as.formula(fixedFormula)
randFormula<-as.formula(randFormula)
# fit asreml 
out<-asreml(fixed = fixedFormula,
            random = randFormula,
            data = MultiTrialTraitData, 
            maxiter = 40, workspace=1000e6, 
            na.method.X="omit")
```

```{r two traits, eval=F}
MultiTrialTraitData<-trialdata %>%
  filter(!is.na(DM),
         !is.na(logFYLD)) %>% 
  mutate(across(c(GID,yearInLoc,
                  CompleteBlocks,
                  IncompleteBlocks,
                  trialInLocYr,
                  repInTrial,
                  blockInRep,
                  TrialType),as.factor)) %>% 
  droplevels

fixedFormula="cbind(DM,logFYLD) ~ yearInLoc*trait + PropNOHAV*trait"
randFormula=paste0("~us(trait,init=c(0,0,0)):GID + idv(trialInLocYr) + at(CompleteBlocks,'Yes'):repInTrial ",
                   "+ at(IncompleteBlocks,'Yes'):blockInRep")
errFormula=paste0("~units:us(trait,init=c(0,0,0))")
require(asreml); 
fixedFormula<-as.formula(fixedFormula)
randFormula<-as.formula(randFormula)
errFormula<-as.formula(errFormula)
# fit asreml 
out<-asreml(fixed = fixedFormula,
            random = randFormula,
            rcov = errFormula,
            data = MultiTrialTraitData, 
            maxiter = 40, workspace=1000e6)
```

```{r two traits heterog error, eval=F}
fixedFormula="cbind(DM,logFYLD) ~ yearInLoc*trait + PropNOHAV*trait"
randFormula=paste0("~us(trait,init=c(0,0,0)):GID + idv(trialInLocYr) + at(CompleteBlocks,'Yes'):repInTrial ",
                   "+ at(IncompleteBlocks,'Yes'):blockInRep")
errFormula=paste0("~units:at(TrialType):us(trait,init=c(0,0,0))")
require(asreml); 
fixedFormula<-as.formula(fixedFormula)
randFormula<-as.formula(randFormula)
errFormula<-as.formula(errFormula)
# fit asreml 
out<-asreml(fixed = fixedFormula,
            random = randFormula,
            rcov = errFormula,
            data = MultiTrialTraitData %>% arrange(TrialType), 
            maxiter = 40, workspace=1000e6)
```

```{r four traits heterog error, eval=F}
# Check the proportion missing for each trait
trialdata %>% 
  summarize(across(any_of(names(SIwts)),~length(which(is.na(.)))/length(.)))
#   logFYLD     HI    DM   MCMDS logRTNO logDYLD logTOPYLD PLTHT
#     <dbl>  <dbl> <dbl>   <dbl>   <dbl>   <dbl>     <dbl> <dbl>
# 1  0.0166 0.0327 0.101 0.00613  0.0104   0.109    0.0186 0.544
MultiTrialTraitData<-trialdata %>%
  # filter(!is.na(DM),
  #        !is.na(logFYLD)) %>% 
  mutate(across(c(GID,yearInLoc,
                  CompleteBlocks,
                  IncompleteBlocks,
                  trialInLocYr,
                  repInTrial,
                  blockInRep,
                  TrialType),as.factor)) %>% 
  droplevels

fixedFormula="cbind(DM,logFYLD,MCMDS,logTOPYLD) ~ yearInLoc*trait + PropNOHAV*trait"
randFormula=paste0("~us(trait):GID + ",
                   "idv(trialInLocYr) + at(CompleteBlocks,'Yes'):repInTrial ",
                   "+ at(IncompleteBlocks,'Yes'):blockInRep")
errFormula=paste0("~units:at(TrialType):us(trait)")
require(asreml); 
fixedFormula<-as.formula(fixedFormula)
randFormula<-as.formula(randFormula)
errFormula<-as.formula(errFormula)
# fit asreml 
out<-asreml(fixed = fixedFormula,
            random = randFormula,
            rcov = errFormula,
            data = MultiTrialTraitData %>% arrange(TrialType), 
            maxiter = 40, workspace=1000e6)
names(SIwts)
```

### Full multi-trait model

7 traits. Heterogeneous error by TrialType. Skip PLTHT b/c \>50% missing data.

```{r seven traits heterog error - skip PLTHT, eval=F}
# Check the proportion missing for each trait
trialdata %>% 
  summarize(across(any_of(names(SIwts)),~length(which(is.na(.)))/length(.)))
#   logFYLD     HI    DM   MCMDS logRTNO logDYLD logTOPYLD PLTHT
#     <dbl>  <dbl> <dbl>   <dbl>   <dbl>   <dbl>     <dbl> <dbl>
# 1  0.0166 0.0327 0.101 0.00613  0.0104   0.109    0.0186 0.544

# SKIP ONLY PLTHT B/C AMOUNT MISSING...
MultiTrialTraitData<-trialdata %>%
  mutate(across(c(GID,yearInLoc,
                  CompleteBlocks,
                  IncompleteBlocks,
                  trialInLocYr,
                  repInTrial,
                  blockInRep,
                  TrialType),as.factor)) %>% 
  droplevels

fixedFormula="cbind(logFYLD,HI,DM,MCMDS,logRTNO,logDYLD,logTOPYLD) ~ yearInLoc*trait + PropNOHAV*trait"
randFormula=paste0("~us(trait):GID + ",
                   "idv(trialInLocYr) + at(CompleteBlocks,'Yes'):repInTrial ",
                   "+ at(IncompleteBlocks,'Yes'):blockInRep")
errFormula=paste0("~units:at(TrialType):us(trait)")
require(asreml); 
fixedFormula<-as.formula(fixedFormula)
randFormula<-as.formula(randFormula)
errFormula<-as.formula(errFormula)
# fit asreml 
out<-asreml(fixed = fixedFormula,
            random = randFormula,
            rcov = errFormula,
            data = MultiTrialTraitData %>% arrange(TrialType), 
            maxiter = 60, workspace=1000e6)
saveRDS(out,file=here::here("output","estimateErrorCov_byTrialType_asreml_2021Aug25.rds"))
as_summary<-summary(out)
saveRDS(as_summary,file=here::here("output","estimateErrorCov_byTrialType_asreml_summary_2021Aug25.rds"))
# 
# ASReml: Wed Aug 25 16:48:25 2021
# 
#  US matrix updates modified 7 times to remain positive definite.
#      LogLik         S2      DF      wall     cpu
# -552132.9648      1.0000471133  16:50:44   107.3 (199 restrained)
#  Notice: NonPosDef US matrix modified
# -172985.8107      1.0000471133  16:52:02    77.5 (199 restrained)
#  -11758.3924      1.0000471133  16:53:15    73.2 (199 restrained)
#   86902.0709      1.0000471133  16:54:31    76.4 (199 restrained)
#  171214.2253      1.0000471133  16:55:47    75.0 (199 restrained)
#  225677.4607      1.0000471133  16:56:57    70.1 (155 restrained)
#  260884.6336      1.0000471133  16:58:06    68.7 (148 restrained)
#  280487.3333      1.0000471133  16:59:13    67.2 (87 restrained)
#  295260.2919      1.0000471133  17:00:20    66.5 (32 restrained)
#  302995.2011      1.0000471133  17:01:30    70.4 (5 restrained)
#  311088.9518      1.0000471133  17:02:46    75.9 (4 restrained)
#  Notice: NonPosDef US matrix modified
#  316645.9289      1.0000471133  17:03:59    72.2 (5 restrained)
#  324430.0701      1.0000471133  17:05:06    67.1 (2 restrained)
#  331602.8926      1.0000471133  17:06:14    67.9
#  335069.2210      1.0000471133  17:07:27    72.8
#  337425.1130      1.0000471133  17:08:50    77.6
#  339834.7415      1.0000471133  17:10:49    98.5
#  340877.5399      1.0000471133  17:13:12   114.1
#  341423.1058      1.0000471133  17:14:50    84.5
#  341734.7595      1.0000471133  17:16:00    70.2
#  341939.2189      1.0000471133  17:17:12    71.2
#  342091.3023      1.0000471133  17:18:27    75.2
#  342215.4180      1.0000471133  17:19:49    81.6
#  342322.6216      1.0000471133  17:21:38    94.1
#  342418.1298      1.0000471133  17:23:04    84.4
#  342504.5682      1.0000471133  17:24:26    79.1
#  342583.3893      1.0000471133  17:25:54    83.8
#  342655.5189      1.0000471133  17:27:22    84.5
#  342721.6172      1.0000471133  17:28:59    85.1
#  342782.2099      1.0000471133  17:30:46    90.1
#  342837.7490      1.0000471133  17:32:17    76.1
#  342888.6246      1.0000471133  17:33:23    65.9
#  342935.2019      1.0000471133  17:34:25    62.7
#  342977.8038      1.0000471133  17:35:29    63.7
#  343016.7322      1.0000471133  17:36:35    65.5
#  343052.2701      1.0000471133  17:37:34    59.8
#  343084.6751      1.0000471133  17:38:37    63.0
#  343114.1894      1.0000471133  17:39:42    64.2
#  343141.0370      1.0000471133  17:40:46    64.7
#  343165.4329      1.0000471133  17:41:46    59.9
#  343187.5770      1.0000471133  17:42:51    64.7
#  343207.6489      1.0000471133  17:43:54    63.5
#  343225.8251      1.0000471133  17:44:58    63.8
#  343242.2660      1.0000471133  17:46:06    67.9
#  343257.1194      1.0000471133  17:47:25    77.2
#  343270.5287      1.0000471133  17:48:49    76.7
#  343282.6168      1.0000471133  17:50:12    81.2
#  343293.5105      1.0000471133  17:51:18    66.5
#  343303.3166      1.0000471133  17:52:25    66.2
#  343312.1366      1.0000471133  17:53:44    75.7
#  343320.0631      1.0000471133  17:55:06    80.0
#  343327.1802      1.0000471133  17:56:25    74.7
#  343333.5731      1.0000471133  17:57:53    79.7
#  343339.3063      1.0000471133  17:59:17    78.3
#  343344.4459      1.0000471133  18:01:01    76.5
#  343349.0523      1.0000471133  18:02:05    64.3
#  343353.1798      1.0000471133  18:03:08    63.0
#  343356.8724      1.0000471133  18:04:13    64.6
#  343360.1798      1.0000471133  18:05:19    66.5
#  343363.1377      1.0000471133  18:06:29    70.0
# US variance structures were modified in 51 instances to make them positive definite
# 
# Finished on: Wed Aug 25 18:06:31 2021
# 
# LogLikelihood not converged


```

Independent but heterogeneous error covariances were fit by TrialType. After 40 iterations (\~30 minutes), likelihood wasn't converged... but looks on the way.

### Result

```{r}
asfit<-readRDS(file=here::here("output","estimateErrorCov_byTrialType_asreml_2021Aug25.rds"))
as_summary<-readRDS(file=here::here("output","estimateErrorCov_byTrialType_asreml_summary_2021Aug25.rds"))
varcomps<-as_summary$varcomp %>% 
  rownames_to_column(var = "VarComp") %>% 
  select(VarComp,component)

# SELECTION INDEX WEIGHTS
SIwts<-c(logFYLD=20,
         HI=10,
         DM=15,
         MCMDS=-10,
         logRTNO=12,
         logDYLD=20,
         logTOPYLD=15)
#         PLTHT=10) 
adjSIwts<-trialdata %>% 
  summarize(across(all_of(names(SIwts)),~sqrt(var(.,na.rm=T)))) %>% 
  magrittr::divide_by(SIwts,.) %>% 
  as.numeric() %>% 
  `names<-`(.,names(SIwts))
```

SIwts, specified by breeder as relative importances, essentially.

Try adjusting weights by dividing by Trait Std. Devs.

Trait Standard Deviations:

```{r}
trialdata %>%
  summarize(across(all_of(names(SIwts)),~sqrt(var(.,na.rm=T)),.names = paste0("sd_","{.col}")));
```

Adjusted Weights:

```{r}
adjSIwts
```

```{r error with orig SIwts}
errorVars<-varcomps %>% 
  filter(grepl("TrialType",VarComp),
         grepl("!trait.",VarComp)) %>% 
  separate(VarComp,c("TrialType","VarParam"),"!trait.",remove = T) %>% 
  separate(VarParam,c("Trait1","Trait2"),":",remove = T) %>% 
  mutate(TrialType=gsub("TrialType_","",TrialType)) %>% 
  nest(VarEsts=c(Trait1,Trait2,component))
errorVars %<>% 
  mutate(siErrorVarEst=map_dbl(VarEsts,
                           function(VarEsts,SIwts){
                             covMat<-VarEsts %>% 
                               spread(Trait2,component) %>% 
                               column_to_rownames("Trait1") %>% 
                               as.matrix() %>% 
                               .[names(SIwts),names(SIwts)]
                             covMat[upper.tri(covMat)]<-t(covMat)[upper.tri(covMat)]
                             siError<-SIwts%*%covMat%*%SIwts
                             return(siError) },SIwts=SIwts))
```

```{r siError vs. plotArea - orig SIwts}
errorVars %>% 
  left_join(trialdata %>% 
              distinct(TrialType,plotArea)) %>% 
  arrange(plotArea) %>% 
  ggplot(.,aes(x=plotArea,y=siErrorVarEst)) + geom_line() + theme_bw() + 
  geom_label(aes(label=TrialType,color=TrialType),size=4) + 
  labs(title = "Unadjusted SIwts - siError vs. plotArea")
```

```{r error with adjSIwts}
adjErrorVars<-varcomps %>% 
  filter(grepl("TrialType",VarComp),
         grepl("!trait.",VarComp)) %>% 
  separate(VarComp,c("TrialType","VarParam"),"!trait.",remove = T) %>% 
  separate(VarParam,c("Trait1","Trait2"),":",remove = T) %>% 
  mutate(TrialType=gsub("TrialType_","",TrialType)) %>% 
  nest(VarEsts=c(Trait1,Trait2,component))
adjErrorVars %<>% 
  mutate(siErrorVarEst=map_dbl(VarEsts,
                           function(VarEsts,SIwts){
                             covMat<-VarEsts %>% 
                               spread(Trait2,component) %>% 
                               column_to_rownames("Trait1") %>% 
                               as.matrix() %>% 
                               .[names(SIwts),names(SIwts)]
                             covMat[upper.tri(covMat)]<-t(covMat)[upper.tri(covMat)]
                             siError<-SIwts%*%covMat%*%SIwts
                             return(siError) },SIwts=adjSIwts))
```

```{r siError vs. plotArea - adjSIwts}
adjErrorVars %>% 
  left_join(trialdata %>% 
              distinct(TrialType,plotArea)) %>% 
  arrange(plotArea) %>% 
  ggplot(.,aes(x=plotArea,y=siErrorVarEst)) + geom_line() + theme_bw() + 
  geom_label(aes(label=TrialType,color=TrialType),size=4) + 
  labs(title = "Adjusted SIwts - siError vs. plotArea")
```

### Convergence check

```{r look at convergence monitor}
asfit$monitor["loglik",] %>% 
  as.numeric %>% plot(.,ylab='loglik',xlab='iter')
```

```{r}
asfit$monitor %>% 
  as.data.frame %>% 
  rownames_to_column("param") %>% 
  select(-constraint) %>% 
  pivot_longer(cols=-c("param"),names_to="iter",values_to="value") %>% 
  mutate(iter=as.numeric(iter)) %>% 
  filter(grepl("!trait",param)) %>% 
  mutate(Component=ifelse(grepl("GID!trait",param),"genVar","errorVar")) %>% 
  ggplot(.,aes(x=iter,y=value,group=param,color=Component)) + geom_line(alpha=0.7) + 
  labs(title="ErrorVarCovars")#geom_label(aes(label=param))
```

```{r}
asfit$monitor %>% 
  as.data.frame %>% 
  rownames_to_column("param") %>% 
  select(-constraint) %>% 
  pivot_longer(cols=-c("param"),names_to="iter",values_to="value") %>% 
  mutate(iter=as.numeric(iter)) %>% 
  filter(!grepl("!trait|!variance|loglik|S2|df",param)) %>% 
  ggplot(.,aes(x=iter,y=value,group=param,color=param)) + geom_line()
```

```{r, fig.width=10}
asfit$monitor %>% 
  as.data.frame %>% 
  rownames_to_column("param") %>% 
  select(-constraint) %>% 
  pivot_longer(cols=-c("param"),names_to="iter",values_to="value") %>% 
  mutate(iter=as.numeric(iter)) %>% 
  filter(grepl("!trait",param)) %>% 
  mutate(Component=ifelse(grepl("GID!trait",param),"genVar","errorVar")) %>% 
  filter(Component=="genVar") %>% 
  ggplot(.,aes(x=iter,y=value,group=param,color=param)) + geom_line(alpha=0.7) + 
  labs(title="Genetic VarCovars") + theme(legend.position = 'bottom')
```

## Approach 2: indirect - regress SI GETGV on Trial-specific SI BLUP

This will be a revised version of the original version of this approach, which is [documented here](https://wolfemd.github.io/IITA_2021GS/inputsForSimulation.html).

**Changes:**

-   Analyze the same subset of trials chosen in Approach 1 above, which is more restrictively chosen than the previous analysis
-   Adopt a SI weighting scheme for computing the Trial-specific SI with BLUPs, which is invariant to which traits are observed
-   Categorize trials according to `plotArea` and/or `TrialType`

**Procedure:**

-   Use the *SELECTION INDEX* GETGV from genomic prediction using all entire available training population and all of the latest available data as a best estimate of "true" net merit

    -   The predictions documented [here](https://wolfemd.github.io/IITA_2021GS/06-GenomicPredictions.html) is the most up-to-date. See the summary of genomic prediction results [here](https://wolfemd.github.io/IITA_2021GS/07-Results.html#Genomic_Predictions).

    -   Alternative best estimate of "true" net merit: actual national performance trial data (e.g. from NCRPs in Nigeria?)

-   For each trial, analyze the cleaned plot-basis data:

    1.  Fit a univariate mixed-model to each trait scored

    2.  Extract trial-specific BLUPs for whatever clones were present

    3.  Compute the SELIND for the current trial using BLUPs for whatever component traits were scored ($SI_{TrialBLUP}$).

        -   Scale weights so their sum always equals the sum of the full set of weights, even when only subsets of traits are present. 

    4.  Regress $SI_{GETGV}$ on the $SI_{TrialBLUP}$

    5.  Extract the $\hat{\sigma}^2_e$ of the regression as the trial-specific estimate of the selection error

-   Use mean (or median) $\hat{\sigma}^2_e$ for each TrialType / plotArea as potential simulation input. Consider a weighted mean/median according to number of clones available to measure $\hat{\sigma}^2_e$ for each trial.

```{bash, eval=F}
screen;
cd ~/IITA_2021GS/;
salloc -n 20 --mem=60G --time=06:00:00;
#export PATH=/programs/R-4.0.5clean-p/bin:$PATH
#export OMP_NUM_THREADS=8
R;
```

```{r inputs for approach 2}
library(genomicMateSelectR); 
library(tidyverse)

# CLEANED PLOT-LEVEL TRIAL DATA
dbdata<-readRDS(here::here("output","IITA_ExptDesignsDetected_2021Aug08.rds"))
# SELECTION INDEX WEIGHTS
SIwts<-c(logFYLD=20,
         HI=10,
         DM=15,
         MCMDS=-10,
         logRTNO=12,
         logDYLD=20,
         logTOPYLD=15) # ,PLTHT=10) 

# FILTER TRIALS TO-BE-CONSIDERED
### Restrict consideration to >2012 
### to measure the selection error during the current "era" at IITA.
### Only trials with >=50% genotyped and key TrialTypes
### Keep only trials with full plotWidth and plotLength meta-data
### Calc plotArea=plotWidth*plotLength (in meters-squared)
trialdata<-dbdata %>% 
  filter(studyYear>=2013,
         !is.na(MaxNOHAV),
         !is.na(plotWidth),
         !is.na(plotLength),
         !is.na(PropNOHAV)) %>% 
  nest(TrialData=-c(studyName,TrialType,plotWidth,plotLength,CompleteBlocks,IncompleteBlocks,MaxNOHAV)) %>% 
  mutate(propGenotyped=map_dbl(TrialData,
                              ~length(which(!is.na(unique(.$FullSampleName))))/length(unique(.$GID))),
         plotArea=plotWidth*plotLength,
         IncompleteBlocks=ifelse(IncompleteBlocks==TRUE,"Yes","No"),
         CompleteBlocks=ifelse(CompleteBlocks==TRUE,"Yes","No")) %>% 
  filter(propGenotyped>=0.5,
         TrialType %in% c("GeneticGain","CET","ExpCET","PYT","AYT","UYT")) %>%
  unnest(TrialData) %>% 
  select(yearInLoc,studyName,studyYear,locationName,TrialType,
         plotArea,plotWidth,plotLength,
         CompleteBlocks,IncompleteBlocks,observationUnitDbId,
         GID,trialInLocYr,repInTrial,blockInRep,PropNOHAV,MaxNOHAV,
         all_of(names(SIwts)))
trialdata %<>% 
  semi_join(trialdata %>% 
              distinct(TrialType,studyName,plotArea,plotWidth,plotLength) %>% 
              group_by(TrialType) %>% 
              summarize(plotArea=median(plotArea)) %>% ungroup())

### ADJUSTED SELIND WEIGHTS
adjSIwts<-trialdata %>% 
  summarize(across(all_of(names(SIwts)),~sqrt(var(.,na.rm=T)))) %>% 
  magrittr::divide_by(SIwts,.) %>% 
  as.numeric() %>% 
  `names<-`(.,names(SIwts))

# SELIND GETGVS (for input to estimateSelectionError func below)
gpreds<-readRDS(file = here::here("output","genomicPredictions_full_set_2021Aug09.rds"))
getgvs<-gpreds$gblups[[1]] %>% 
  filter(predOf=="GETGV") %>% 
  select(GID,SELIND,all_of(names(SIwts)))
siadj_getgvs<-getgvs %>% 
  mutate(SELIND=as.numeric(getgvs %>% 
                             select(-SELIND,-GID) %>% 
                             as.matrix(.)%*%adjSIwts)) %>% 
  select(GID,SELIND)
si_getgvs<-getgvs %>% 
  select(GID,SELIND)

trialdata %<>% 
  nest(TrialData=-c(studyYear,locationName,studyName,TrialType,CompleteBlocks,IncompleteBlocks,MaxNOHAV,plotArea))

# SOURCE FUNCTION estimateSelectionError() 
source(here::here("code","estimateSelectionError.R"))
```

```{r unit test inputs for estimateSelectionError, eval=F}
###### unit test inputs for estimateSelectionError
# TrialData<-trialdata$TrialData[[1]]
# CompleteBlocks<-trialdata$CompleteBlocks[[1]]
# IncompleteBlocks<-trialdata$IncompleteBlocks[[1]]
# getgvs<-si_getgvs
# TrialData<-trialdata %>% filter(propGenotyped>0.75) %>% slice(4) %$% TrialData[[1]]
# CompleteBlocks<-trialdata %>% filter(propGenotyped>0.75) %>% slice(4) %$% CompleteBlocks[[1]]
# IncompleteBlocks<-trialdata %>% filter(propGenotyped>0.75) %>% slice(4) %$% IncompleteBlocks[[1]]
# ncores=4
# rm(TrialData,CompleteBlocks,IncompleteBlocks)
```

Run function `estimateSelectionError()` across trials to estimation selection errors.

```{r, eval=F}
# ORIGINAL SIwts
require(furrr); plan(multisession, workers = 20)
options(future.globals.maxSize=+Inf); options(future.rng.onMisuse="ignore")
trialdata %<>% 
  mutate(SelectionError=future_pmap(.,estimateSelectionError,
                                    SIwts=SIwts,getgvs=si_getgvs))
plan(sequential)
saveRDS(trialdata,here::here("output","estimateSelectionError_origSIwts_2021Aug24.rds"))

# ADJUSTED SIwts
require(furrr); plan(multisession, workers = 20)
options(future.globals.maxSize=+Inf); options(future.rng.onMisuse="ignore")
trialdata %<>% 
  mutate(SelectionError=future_pmap(.,estimateSelectionError,
                                    SIwts=adjSIwts,getgvs=siadj_getgvs))
plan(sequential)
saveRDS(trialdata,here::here("output","estimateSelectionError_adjSIwts_2021Aug24.rds"))

```

# Conclusions

The analysis seems far from conclusive. Do we even expect the relation b/t plot-size and error variance to be linearly decreasing?

Certainly this will vary by breeding, location, and much more.

There are also possibly many alternative methods to analyze this problem, and alternative datasets to choose.

**NEW APPROACH:** While inconclusive, this exercise emphasizes a key concern for conducting simulations that alter the VDP: that the cost-benefit balance could depend on the relative information value/selection accuracy/error variance of different plot sizes and trial configurations. As a result, I propose to simulate a range of error-vs-plot size as part of the [baseline simulation](https://wolfemd.github.io/BreedingSchemeOpt/baselineSim.html) of each breeding program. ***If*** we observe a shift-point in the cost-benefit analysis we can then work with breeding programs to determine where their data indicate they lie and what changes are subsequently recommended.
